---
title: "Reno_Ames_EDA"
author: "James Reno"
date: '2022-05-17'
output: html_document
---

# EDA graphs for presentation and Shiny App

I performed my EDA in Python, but reproduced the visualizations in R to match the rest of my team.

These are the final graphs I used in the project presentation.

```{r}

library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
theme_set(theme_bw())
library(MASS)


```


```{r}

ames_X = read.csv("ML-Ames/data/Ames_X.csv")
ames_Y = read.csv("ML-Ames/data/Ames_Y.csv")

colnames(ames_X)
ames_X = ames_X[,-1]
ames_Y = ames_Y[,-1]

ames_raw = read.csv("Ames_HousePrice.csv")
```


```{r graph of quality times square feet versus sale price}

ames = cbind(ames_X,ames_Y)

colnames(ames)

g = ames %>% mutate(SalePriceShort = round(ames_Y/1000),2) %>% 
  ggplot(aes(x=QualXSF, y=SalePriceShort, size=GrLivArea, color=AgeSold)) +
  geom_point(alpha=0.1)  +
  labs(x="Quality * Square Feet",y="Sale Price\n($1,000 USD)", title="Quality * Square Feet vs Sale Price",  size="Square Feet",color="Age")  +
  scale_color_continuous(trans = 'reverse')  + scale_color_gradient(low="#534666", high="#EEB462") +
  scale_x_continuous(labels = scales::comma) + theme_bw() + scale_size_continuous(labels = scales::comma) + geom_smooth(method=lm, se=FALSE, show.legend = FALSE, color = '#534666') + theme_ipsum(base_size=16)
 
ggsave(
  filename="QualitySF_Price.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)
#+ guides(colour = guide_legend(override.aes = list(alpha = 0)))
ames$SalePrice
```

```{r scatter quality by age}
g=ames %>% 
  ggplot(aes(x=as.factor(OverallQual), y=AgeSold)) +
  geom_boxplot()  +
  labs(x="Quality",y="Age", title="Quality vs Age")  + 
  theme_bw() + theme_ipsum(base_size=16)

ggsave(
  filename="Quality_by_Age.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)
```

```{r square feet by age}
g = ames_raw %>% mutate(AgeSold = YrSold - YearBuilt) %>% 
  ggplot(aes(x=GrLivArea, y=AgeSold)) +
  geom_point()  +
  labs(x="Square Feet",y="Age", title="Square Feet vs Age")  + 
  theme_bw() + theme_ipsum(base_size=16) + scale_x_continuous(labels = scales::comma)

ggsave(
  filename="SF_by_Age.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)
```


```{r footprint vs price}
ames %>% mutate(SalePriceLog = log(ames_Y)) %>% 
  ggplot(aes(x=house_footprint, y=SalePriceLog, size=GrLivArea, color=AgeSold)) +
  geom_point(alpha=0.1)  +
  labs(x="House Footprint",y="Log(Sale Price)",  size="Square Feet",color="Age")  +
  scale_color_continuous(trans = 'reverse')  + scale_color_gradient(low="#534666", high="#EEB462") +
  scale_x_continuous(labels = scales::comma) + theme_bw() + scale_size_continuous(labels = scales::comma) + geom_smooth(method=lm, se=FALSE, show.legend = FALSE, color = '#534666') + theme_ipsum(base_size=16)
```


## Not used.

These three plots were made to illustrate examples of positive, negative, and no relationship in the data as an introduction for realtors who may not know what types of relationships a data team is looking for. Team chose not to include.


```{r positive relationship example}

x_vals = seq(0,5,.01)
err=seq(-3,3,.5)
err2 = rnorm(1000)
err3 = sample(err2,1000,replace=FALSE)

x_pos = sample(x_vals,1000,replace=TRUE)
y_pos = (x_pos*1)+err3

x_y_pos2 = data.frame(x_neg,y_neg)



g = x_y_pos2  %>% ggplot(aes(x=x_pos, y=y_pos)) +
  geom_point(alpha=0.1)  +
  labs(x="Feature 1",y="Feature 2",title="Positive Relationship") +
  geom_smooth(method=lm, se=FALSE, show.legend = FALSE) + theme_ipsum(base_size=16) + 
  scale_y_continuous(breaks=seq(1,6,by=1))


ggsave(
  filename="Positive_rel.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)

```



```{r negative relationship example}

x_vals = seq(0,5,.01)
err=seq(-3,3,.5)
err2 = rnorm(1000)
err3 = sample(err2,1000,replace=FALSE)

x_neg = sample(x_vals,1000,replace=TRUE)
y_neg = 6+(x_neg*-1)+err3

x_y_neg2 = data.frame(x_neg,y_neg)



g = x_y_neg2  %>% ggplot(aes(x=x_neg, y=y_neg)) +
  geom_point(alpha=0.1)  +
  labs(x="Feature 1",y="Feature 2",title="Negative Relationship") +
  geom_smooth(method=lm, se=FALSE, show.legend = FALSE) + theme_ipsum(base_size=16) + 
  scale_y_continuous(breaks=seq(1,6,by=1))

ggsave(
  filename="Negative_rel.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)

```


```{r no relationship example}

x_vals = seq(0,5,.01)
err=seq(-3,3,.5)
err2 = rnorm(1000)
err3 = sample(err2,1000,replace=FALSE)

x_no = sample(x_vals,1000,replace=TRUE)
y_no = 2.5+err3

x_y_no2 = data.frame(x_neg,y_neg)



g=x_y_no2  %>% ggplot(aes(x=x_no, y=y_no)) +
  geom_point(alpha=0.1)  +
  labs(x="Feature 1",y="Feature 2",title="No Relationship") +
  geom_smooth(method=lm, se=FALSE, show.legend = FALSE) + theme_ipsum(base_size=16) + 
  scale_y_continuous(breaks=seq(1,6,by=1))

ggsave(
  filename="No_rel.jpeg",
  plot = g,
  width = 8.5 - 2,
  height = (11 - 2) / 2,
  units = "in",
  dpi = 300
)

```

